\textls[-15]{\thetitle \mbox{(Technical Report)}}

### 2022-06-30 syntax.tex
resolve & add comments

\begin{figure}[!htb]
  \begin{align*}
    &\Gamma~:~"Con"&&\mathboxtext{\goodCtx{\Gamma}{}} &&T~:~"Ty"~\Gamma&&\mathboxtext{\goodType{\Gamma}{T}{}}  &&\gamma~:~"Tms"~\Gamma~\Delta&&\mathboxtext{\goodSub{\Gamma}{\gamma}{\Delta}}  \\ & t~:~"Tm"~\Gamma~T&&\mathboxtext{\goodTerm{\Gamma}{t}{T}} 
    &&\sigma~:~"Sig"~\Gamma~n&&\mathboxtext{\goodSig{\Gamma}{\sigma}{n}}  &&\tau~:~"WSig"~\Gamma~n&&\mathboxtext{\goodWSig{\Gamma}{\tau}{n}} \\ &h~:~"Inh"~\Gamma~\sigma_1~\sigma_2&&\mathboxtext{\goodInh{\Gamma}{h}{\sigma}{\tau}} && f~:~"Seal"~\Gamma~\sigma~T&&\mathboxtext{\goodSeal{\Gamma}{f}{\sigma}{T}}\YZ{A little unfriendly to the reader that there are three different things that sigma can range over. Besides, later you also use gamma to range over substitutions and s to range over W-types, which add to the confusion.}\EDJreply{Fair point. I will fix sigma and tau to be about signatures and W signature, and use gamma as substitutions}
  \end{align*}
\caption{Judgement forms, as Agda types and as presented in this section.}
\end{figure}

@@

Sealing judgement is still about the important term $h$ we mentioned
earlier. Thanks to $\cC$, a newly added term $t_{new}$\YZ{By 'a newly
added term', do you mean a term that follows field 'a'?} is now in the
context containing $\cC\{a : \bW \sigma; ..\}$.
But if we want $t_{new}$ to be inherited(reused) when $\sigma$ is
extended, we would like $t_{new}$ to be in the context containing
$\cC\{a : \cU; ..\}$ instead. The sealing is responsible for this
because we have a sealing $\goodTerm{\Gamma, \cC\{a : \bW \sigma;
..\}}{s}{\cC\{a : \cU; ..\}}$ = $\goodTerm{\Gamma, (\Sigma a : \bW
\sigma,..)}{s}{(\Sigma a : \cU,..)}$. Then we can construct
$t_{new}$ in the context containing $(\Sigma a : \cU,..)$, ready for
inheritance.
Apparently $t_{new}$ in this case cannot be doing \textit{exhaustiveness
checking}, but it can freely use any \textit{exhaustiveness-checked}
recursor.\YZ{It occurs to me that you use the word "sealing" to mean
different (though tangentially related) things. Previously you used it
as in "sealed families".}\EDJreply{They are actually the same thing.
This sealing in metatheory is much stronger than that of the implemented
plugin. Downside of the strong sealing is that if the user really use
this version of sealing, the user will have to annotate what the sealed
family type is. Now in the plugin, the sealed family type is almost
fixed (directly inferred using a fixed rules). So basically this
complete version of sealing (in metatheory) can deduce that of seal (in
plugin implementation) }

@@

$$
\Rule[name=Hdler]
{\goodType{\Gamma}{A}{}
\quad \goodType{\Gamma, A}{B}{}
\quad \goodType{\Gamma}{R}{}
}
{\goodType{\Gamma}{"RecSig1"~A~B~R \equiv (\Pi A (\Pi (\Pi B (R[\pi_1\circ\pi_1])) R[\pi_1\circ\pi_1]) )}{}\YZ{What is R[pi1 circ pi1]?}\EDJreply{It is the composition of the substitution. Thanks for pointing out now I have included circ into the typesetting} }
\quad 
\Rule[name=Hdlers]
{\goodWSig{\Gamma}{\sigma}{n}
\quad \goodType{\Gamma}{R}{}
}
{\goodSig{\Gamma}{"RecSig"~\sigma~R}{}}
$$
$$
\Rule[name=Hdler-Proj]
{ \goodWSig{\Gamma}{\sigma}{N}
\quad \goodTerm{\Gamma}{o}{\cL("RecSig"~\sigma~R)}
\quad n < N
}
{\goodTerm{\Gamma}{"prjR"~n~o}{("RecSig1"~(\pi_1^n \sigma)~(\pi_2^n \sigma)~R)[\pi_1]}}
$$
$$
\Rule[name=Rec-Constr]
{ \goodWSig{\Gamma}{\sigma}{N}
\quad \goodType{\Gamma}{T}{\bW \sigma}
\quad \goodType{\Gamma}{R}{}
\quad \goodTerm{\Gamma}{r}{\cL("RecSig"~\sigma~R)}
\quad \goodTerm{\Gamma}{w}{El~("W"~T)}
}
{\goodTerm{\Gamma}{"Wrec"~r~w}{R}}
$$

@@

Intuitively \ruleref{Hdler} is indicating the type of recursor handler
for one specific constructor, and \ruleref{Hdlers} is the linkage type
of the bundle of all recursor handlers of one given inductive type. We
omit the definition of "RecSig" but \ruleref{Hdler-Proj} indicates its
(apparent) internal definition. "prjR" is also the handler selection
from the handler linkage, and is helpful for the formulation of
$\beta$-rule of "Wrec" (omitted here). Finally \ruleref{Rec-Constr} is
the used to construct a concrete recursor given a specific linkage of
handlers. All these rules satisfy substitution laws and omitted here.
\YZ{Seems like you omit many things here, which causes hardship for the reader (at least for me!). I think it's better to list all the rules and decide later what to throw to the appendix.}\EDJreply{Do you mean you want to see substitution laws and beta and eta laws? They are just commutating as expected I can write down them if you really want. And what else do you think is helpful}

Again, we still use \textit{decoupling} of the syntax of the case handlers and exhaustiveness checking to resolve \ref{chg:extensible-inductive-type}. The linkage is the part of defining case handlers and thus inheritable like other family.
While any fields using "Wsup" and "Wrec" directly relies on the
$\bW(\sigma)$ (can be spotted from \ruleref{Ind-Term} and
\ruleref{Rec-Constr}) and thus non-inheritable---in these cases, "Wsup"
can be understood as concrete constructors and "Wrec" are
\textit{exhaustiveness checking}. These fields together with inductive
types are the ones sealing is responsible for hiding. After compilation
of linkage types into existential types, we can easily hide the
\textit{concrete} definition of the inductive type while exposing all
the fields using "Wsup" and "Wrec" like an abstract interface.
Apparently, future fields relying on this abstract interface will have no
problem on inheritance.
Of course, since "Wsup" and "Wrec" directly rely on the concrete inductive definition, we will have to override these fields whenever inductive type is overridden. For "Wsup" we only need to override each constructor and for "Wrec" it is just doing exhaustiveness checking inside each children family. The recursor handlers however, can be safely inherited.
\YZ{I suppose the terminology 'sealing' is a reference to ML modules? In
ML, sealing has the connotation of hiding things behind an existential
type. So using the same word may cause the reader to have false
presumptions about what is going on here.}\EDJreply{You are right. Our "Sealing" is not hiding a particular field into an existential type, but our seal is make several fields together into an "existential type". Basically it hides the concrete definition of inductive type so that future fields doesn't have to be defined upon that.}\EDJreply{I will resolve this comment once I add the throughout example and you find it clear---because I haven't changed any text}

There are also cases we don't need to seal anything (i.e., for doing
exhaustiveness checking when using recursor), and thus the "Seal-Id" is
used. For notation clarity, we will sometimes use $"sf'"$ to emphasize
that we consider that seal $s$ as a term $("sf'" \ s)$.

>>>>

\begin{figure}[!htb]
  \begin{align*}
    &\Gamma~:~"Con"&&\mathboxtext{\goodCtx{\Gamma}{}} &&T~:~"Ty"~\Gamma&&\mathboxtext{\goodType{\Gamma}{T}{}}  &&\sigma~:~"Tms"~\Gamma~\Delta&&\mathboxtext{\goodSub{\Gamma}{\sigma}{\Delta}}  \\ & t~:~"Tm"~\Gamma~T&&\mathboxtext{\goodTerm{\Gamma}{t}{T}} 
    &&\sigma~:~"Sig"~\Gamma~n&&\mathboxtext{\goodSig{\Gamma}{\sigma}{n}}  &&\sigma~:~"WSig"~\Gamma~n&&\mathboxtext{\goodWSig{\Gamma}{\sigma}{n}} \\ &h~:~"Inh"~\Gamma~\sigma~\tau&&\mathboxtext{\goodInh{\Gamma}{h}{\sigma}{\tau}} && f~:~"Seal"~\Gamma~\sigma~A&&\mathboxtext{\goodSeal{\Gamma}{f}{\sigma}{A}}\YZ{A little unfriendly to the reader that there are three different things that sigma can range over. Besides, later you also use gamma to range over substitutions and s to range over W-types, which add to the confusion.}
  \end{align*}
\caption{Judgement forms, as Agda types and as presented in this section.}
\end{figure}

@@

Sealing judgement is still about the important term $h$ we mentioned
earlier. Thanks to $\cC$, a newly added term $t_{new}$\YZ{By 'a newly
added term', do you mean a term that follows field 'a'?} is now in the
context containing $\cC\{a : \bW \sigma; ..\}$.
But if we want $t_{new}$ to be inherited(reused) when $\sigma$ is
extended, we would like $t_{new}$ to be in the context containing
$\cC\{a : \cU; ..\}$ instead. The sealing is responsible for this
because we have a sealing $\goodTerm{\Gamma, \cC\{a : \bW \sigma;
..\}}{s}{\cC\{a : \cU; ..\}}$ = $\goodTerm{\Gamma, (\Sigma a : \bW
\sigma,..)}{s}{(\Sigma a : \cU,..)}$. Then we can construct
$t_{new}$ in the context containing $(\Sigma a : \cU,..)$, ready for
inheritance.
Apparently $t_{new}$ in this case cannot be doing \textit{exhaustiveness
checking}, but it can freely use any \textit{exhaustiveness-checked}
recursor.
%\YZ{It occurs to me that you use the word "sealing" to mean
%different (though tangentially related) things. Previously you used it
%as in "sealed families".}\EDJreply{They are actually the same thing.
%This sealing in metatheory is much stronger than that of the implemented
%plugin. Downside of the strong sealing is that if the user really use
%this version of sealing, the user will have to annotate what the sealed
%family type is. Now in the plugin, the sealed family type is almost
%fixed (directly inferred using a fixed rules). So basically this
%complete version of sealing (in metatheory) can deduce that of seal (in
%plugin implementation)}

@@

$$
\Rule[name=Hdler]
{\goodType{\Gamma}{A}{}
\quad \goodType{\Gamma, A}{B}{}
\quad \goodType{\Gamma}{R}{}
}
{\goodType{\Gamma}{"RecSig1"~A~B~R \equiv (\Pi A (\Pi (\Pi B (R[\pi_1\circ\pi_1])) R[\pi_1\circ\pi_1]) )}{}}
\quad 
\Rule[name=Hdlers]
{\goodWSig{\Gamma}{\sigma}{n}
\quad \goodType{\Gamma}{R}{}
}
{\goodSig{\Gamma}{"RecSig"~\sigma~R}{}\YZ{Why does this Sig not have a superscript?}}
$$
$$
\Rule[name=Hdler-Proj]
{ \goodWSig{\Gamma}{\sigma}{N}
\quad \goodTerm{\Gamma}{o}{\cL("RecSig"~\sigma~R)}
\quad n < N
}
{\goodTerm{\Gamma}{"prjR"~n~o}{("RecSig1"~(\pi_1^n \sigma)~(\pi_2^n \sigma)~R)[\pi_1]}}
$$
$$
\Rule[name=Rec-Constr]
{ \goodWSig{\Gamma}{\sigma}{N}
\quad \goodType{\Gamma}{T}{\bW \sigma}
\quad \goodType{\Gamma}{R}{}
\quad \goodTerm{\Gamma}{r}{\cL("RecSig"~\sigma~R)}
\quad \goodTerm{\Gamma}{w}{El~("W"~T)}
}
{\goodTerm{\Gamma}{"Wrec"~r~w}{R}}
$$

@@

Intuitively \ruleref{Hdler} is indicating the type of recursor handler
for one specific constructor, and \ruleref{Hdlers} is the linkage type
of the bundle of all recursor handlers of one given inductive type. We
omit the definition of "RecSig" but \ruleref{Hdler-Proj} indicates its
(apparent) internal definition. "prjR" is also the handler selection
from the handler linkage, and is helpful for the formulation of
$\beta$-rule of "Wrec" (omitted here). Finally \ruleref{Rec-Constr} is
used to construct a concrete recursor given a specific linkage of
handlers. All these rules satisfy substitution laws and omitted here.

Again, we still use \textit{decoupling} of the syntax of the case
handlers and exhaustiveness checking to resolve
\ref{chg:extensible-inductive-type}. We use linkages to define
case handlers and thus they are inheritable like other families.
While any field using "Wsup" and "Wrec" directly relies on the
$\bW\sigma$ (can be spotted from \ruleref{Ind-Term} and
\ruleref{Rec-Constr}) and thus non-inheritable---in these cases, "Wsup"
can be understood as concrete constructors and "Wrec" are
\textit{exhaustiveness checking}. These fields together with inductive
types are the ones sealing is responsible for hiding. After compilation
of linkage types into existential types, we can easily hide the
\textit{concrete} definition of the inductive type while exposing all
the fields using "Wsup" and "Wrec" like an abstract interface.
Apparently, future fields relying on this abstract interface will have no
problem on inheritance.
Of course, since "Wsup" and "Wrec" directly rely on the concrete inductive definition, we will have to override these fields whenever inductive type is overridden. For "Wsup" we only need to override each constructor and for "Wrec" it is just doing exhaustiveness checking inside each children family. The recursor handlers however, can be safely inherited.
\YZ{I suppose the terminology 'sealing' is a reference to ML modules? In
ML, sealing has the connotation of hiding things behind an existential
type. So using the same word may cause the reader to have false
presumptions about what is going on here.}\EDJreply{You are right. Our "Sealing" is not hiding a particular field into an existential type, but our seal is make several fields together into an "existential type". Basically it hides the concrete definition of inductive type so that future fields doesn't have to be defined upon that.}\EDJreply{I will resolve this comment once I add the throughout example and you find it clear---because I haven't changed any text}

There are also cases we don't need to seal anything (i.e., for doing
exhaustiveness checking when using recursor), and thus \ruleref{Seal-Id} is
used. For notation clarity, we will sometimes use $"sf'"$ to emphasize
that we consider that seal $s$ as a term $("sf'" \ s)$.\YZ{What is sf and sf'? sf first appears in Compile rules, so explain sf there.}
